### 2017-06-14 connect_year_writer.Rmd
Working on some JOINS

---
title: "Connecting scene direction to year and writer"
author: "Julia Silge"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE, 
                      echo = FALSE, dpi = 360)
options(width=80, dplyr.width = 150)
library(ggplot2)
library(silgelib)
theme_set(theme_roboto())
```

## Reading in the data

```{r pronoun_bigrams}
library(tidyverse)
library(tidytext)
library(stringr)

all_tsvs <- paste0("setDirections/", list.files(path = "setDirections/", pattern = ".tsv$"))

pronoun_bigrams <- all_tsvs %>%
    map_df(~data_frame(lines = read_lines(.x)) %>%
               filter(str_detect(lines, "^[0-9]")) %>%
               separate(lines, c("line", "text"), sep = "\t") %>%
               mutate(scriptID = as.integer(str_extract(.x, "[\\d]+"))) %>%
               unnest_tokens(bigram, text, token = "ngrams", 
                             n = 2, collapse = FALSE) %>%
               separate(bigram, c("word1", "word2"), sep = " ") %>%
               filter(word1 %in% c("he", "she"))) %>%
    select(scriptID, line, word1, word2)

pronoun_bigrams
```

## Joining the bigrams to other metadata about the scripts

```{r pronoun_imdb, dependson="pronoun_bigrams"}

mapping <- read_csv("full_mapping.csv") %>%
    rename(imdb = imdb_id)

gender <- read_csv("gender.csv") %>%
    filter(important == "1",
           gender != "NULL")

genre <- read_tsv("imdb-genre.tsv") %>%
    rename(imdb = imdb_id)

metadata <- read_tsv("imdb-meta-data-title.tsv") %>%
    select(imdb, year) ## could also get title here if we want later

pronoun_imdb <- pronoun_bigrams %>%
    left_join(mapping, by = c("scriptID" = "id")) 
```

What kind of joining can I do? Year, genre, gender of writers/etc

```{r, dependson="pronoun_imdb"}
pronoun_imdb %>%
    inner_join(metadata, by = "imdb") ## to get year

pronoun_imdb %>%
    inner_join(genre, by = "imdb") ## to get genre

pronoun_imdb %>%
    inner_join(gender, by = "imdb") ## to get gender of writers, etc

```