### 2017-06-18 connect_year_writer.Rmd
Add some explanation of modeling

---
title: "Connecting scene direction to year and writer"
author: "Julia Silge"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE, 
                      echo = TRUE, dpi = 360)
options(width=80, dplyr.width = 150)
library(ggplot2)
library(silgelib)
theme_set(theme_roboto())
```

## Reading in the data

```{r pronoun_bigrams}
library(tidyverse)
library(tidytext)
library(stringr)

all_tsvs <- paste0("setDirections/", list.files(path = "setDirections/", pattern = ".tsv$"))

pronoun_bigrams <- all_tsvs %>%
    map_df(~data_frame(lines = read_lines(.x)) %>%
               filter(str_detect(lines, "^[0-9]")) %>%
               separate(lines, c("line", "text"), sep = "\t") %>%
               mutate(scriptID = as.integer(str_extract(.x, "[\\d]+"))) %>%
               unnest_tokens(bigram, text, token = "ngrams", 
                             n = 2, collapse = FALSE) %>%
               separate(bigram, c("word1", "word2"), sep = " ") %>%
               filter(word1 %in% c("he", "she"))) %>%
    select(scriptID, line, word1, word2)

pronoun_bigrams
```

## Joining the bigrams to other metadata about the scripts

```{r pronoun_imdb, dependson="pronoun_bigrams"}

mapping <- read_csv("full_mapping.csv") %>%
    rename(imdb = imdb_id)

gender <- read_csv("gender.csv") %>%
    filter(important == "1",
           gender != "NULL")

genre <- read_tsv("imdb-genre.tsv") %>%
    rename(imdb = imdb_id)

metadata <- read_tsv("imdb-meta-data-title.tsv") %>%
    select(imdb, year) ## could also get title here if we want later

pronoun_imdb <- pronoun_bigrams %>%
    left_join(mapping, by = c("scriptID" = "id")) 
```

What kind of joining can I do? Year, genre, gender of writers/etc

```{r, dependson="pronoun_imdb"}

pronoun_imdb %>%
    inner_join(genre, by = "imdb") ## to get genre

pronoun_imdb %>%
    inner_join(gender, by = "imdb") ## to get gender of writers, etc

```

## Change by year

We can use logistic regression modeling to find the words paired with "she" that have changed the fastest with time, either in a positive or negative direction.

```{r words_by_year, dependson="pronoun_imdb"}
words_by_year <- pronoun_imdb %>%
    inner_join(metadata, by = "imdb") %>% ## to get year
    filter(word1 == "she",
           str_detect(word2, "[a-z]+"),
           word2 != "look",
           nchar(word2) > 2,
           !is.na(year),
           year > 1980) %>%
    mutate(year = 5 * year %/% 5) %>%
    add_count(year) %>%
    rename(year_total = n) %>%
    add_count(word2) %>%
    rename(word_total = n) %>%
    filter(word_total > 50) %>%
    count(word2, year, year_total)

slopes <- words_by_year %>%
    nest(-word2) %>%
    mutate(models = map(data, ~ glm(cbind(n, year_total) ~ year, ., 
                                    family = "binomial"))) %>%
    unnest(map(models, tidy)) %>%
    filter(term == "year") %>%
    arrange(estimate)
```

```{r contractions, dependson="words_by_year", fig.width=7, fig.height=5}
words_by_year %>%
    inner_join(slopes %>%
                   top_n(10, estimate), 
               by = "word2") %>%
    ggplot(aes(year, n / year_total, color = word2)) +
    geom_line(alpha = 0.8, size = 1.3) +
    labs(x = NULL, y = "Word frequency",
         title = "Words paired with 'she' in script set directions",
         subtitle = "Many of the words increasing in frequency with time are negative contractions")
```

```{r increasing, dependson="words_by_year", fig.width=7, fig.height=5}
words_by_year %>%
    inner_join(slopes %>%
                   filter(str_detect(word2, "^[^’]*$")) %>%
                   top_n(10, estimate), 
               by = "word2") %>%
    ggplot(aes(year, n / year_total, color = word2)) +
    geom_line(alpha = 0.8, size = 1.3) +
    labs(x = NULL, y = "Word frequency",
         title = "Words paired with 'she' in script set directions",
         subtitle = "More recently, women are more likely to inspect, drift, steal, exhale,  breathe")
```

```{r decreasing, dependson="words_by_year", fig.width=7, fig.height=5}
words_by_year %>%
    inner_join(slopes %>%
                   top_n(-10, estimate), 
               by = "word2") %>%
    ggplot(aes(year, n / year_total, color = word2)) +
    geom_line(alpha = 0.8, size = 1.3) +
    labs(x = NULL, y = "Word frequency",
         title = "Words paired with 'she' in script set directions",
         subtitle = "In the past, women were more likely to jab, snuggle, fling, whirl, shriek")
```

I'm picturing an interactive version of this chart where the user can switch from increasing to decreasing, hover over the lines to see how they are changing. Maybe decide how many to see?

Let's check the words that change for "he", to compare. (This probably won't go into the final essay.)

```{r words_by_year_he, dependson="pronoun_imdb"}
words_by_year <- pronoun_imdb %>%
    inner_join(metadata, by = "imdb") %>% ## to get year
    filter(word1 == "he",
           str_detect(word2, "[a-z]+"),
           word2 != "look",
           nchar(word2) > 2,
           !is.na(year),
           year > 1980) %>%
    mutate(year = 5 * year %/% 5) %>%
    add_count(year) %>%
    rename(year_total = n) %>%
    add_count(word2) %>%
    rename(word_total = n) %>%
    filter(word_total > 100) %>%
    count(word2, year, year_total)

slopes <- words_by_year %>%
    nest(-word2) %>%
    mutate(models = map(data, ~ glm(cbind(n, year_total) ~ year, ., 
                                    family = "binomial"))) %>%
    unnest(map(models, tidy)) %>%
    filter(term == "year") %>%
    arrange(estimate)

# increasing for "he"
slopes %>% 
    top_n(10, estimate)

# decreasing for "he"
slopes %>% 
    top_n(-10, estimate)
```

So the change in contractions looks like it is more about how set directions are changing as a whole, not a difference for how women are portrayed. For the final modeling by year, I'll plan to filter contractions out.

Also, HA -- women are exhaling and men are inhaling.