### 2017-06-23 connect_year_writer.Rmd
Round out some thoughts on viz ideas, analysis

---
title: "Connecting scene direction to year and writer"
author: "Julia Silge"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE, 
                      echo = TRUE, dpi = 360)
options(width=80, dplyr.width = 150)
library(ggplot2)
library(silgelib)
theme_set(theme_roboto())
```

## Reading in the data

```{r pronoun_bigrams}
library(tidyverse)
library(tidytext)
library(stringr)

all_tsvs <- paste0("setDirections/", list.files(path = "setDirections/", pattern = ".tsv$"))

all_lines <- all_tsvs %>%
    map_df(~data_frame(lines = read_lines(.x)) %>%
               filter(str_detect(lines, "^[0-9]")) %>%
               separate(lines, c("line", "text"), sep = "\t") %>%
               mutate(scriptID = as.integer(str_extract(.x, "[\\d]+"))))

pronoun_bigrams <- all_lines %>%
    unnest_tokens(bigram, text, token = "ngrams", 
                  n = 2, collapse = FALSE) %>%
    separate(bigram, c("word1", "word2"), sep = " ") %>%
    filter(word1 %in% c("he", "she")) %>%
    select(scriptID, line, word1, word2)

pronoun_bigrams
```

## Overall ratios

```{r pronoun_ratio, dependson="pronoun_bigrams", fig.width=7, fig.height=6}
pronoun_ratio <- pronoun_bigrams %>%
    count(word1, word2) %>%
    spread(word1, n, fill = 0) %>%
    mutate(total = he + she,
           he = (he + 1) / sum(he + 1),
           she = (she + 1) / sum(she + 1),
           logratio = log2(she / he)) %>%
    arrange(desc(logratio)) %>%
    filter(!(word2 %in% c("door", "ad", "r", "e", "he", "dead", 
                          "len", "queen", "lp", "room", "other", "two", "ar")))

pronoun_ratio %>%
    filter(total > 100) %>%
    group_by(logratio < 0) %>%
    top_n(15, abs(logratio)) %>%
    ungroup() %>%
    mutate(word = reorder(word2, logratio)) %>%
    ggplot(aes(word, logratio, color = logratio < 0)) +
    geom_segment(aes(x = word, xend = word,
                     y = 0, yend = logratio), 
                 size = 1.1, alpha = 0.6) +
    geom_point(size = 3.5) +
    coord_flip() +
    labs(x = NULL, 
         y = "Relative appearance after 'she' compared to 'he'",
         title = "Words paired with 'he' and 'she' in film set directions",
         subtitle = "From 3000+ films, verbs mentioned 100+ times") +
    scale_color_discrete(name = "", labels = c("More 'she'", "More 'he'")) +
    scale_y_continuous(breaks = seq(-3, 3),
                       labels = c("0.125x", "0.25x", "0.5x", 
                                  "Same", "2x", "4x", "8x"))
```

Overall, we see that women are more likely to squeal, giggle and sob in these set directions, while men are more likely to shoot, kill, vault, and amble. With the particular filtering and cleaning used to make this plot, there are `r pronoun_ratio %>% filter(total > 100) %>% nrow()` words in the data set that we can measure a ratio for, from over 4 times for "squeals" on down. One option for a visualization would be something like [this](https://qz.com/836813/10000-words-ranked-according-to-trumpiness/) from Quartz. We could put all 800+ words into the visualization as circles, similar to that type of layout. We could label particularly interesting ones, as well as all the ones at the extremes.

What types of words are equally likely to be paired with "he" and "she"?

```{r dependson="pronoun_ratio"}
pronoun_ratio %>%
    filter(total > 100) %>%
    arrange(abs(logratio))
```

Both men and women concentrate, prop things/themselves, finish, and "just" or "might" do something.

Another interesting layer to add somehow would be some snippets of text to the visualization. For example, for "squeals", we can see some of the examples.

```{r dependson="pronoun_ratio"}
all_lines %>%
    filter(str_detect(text, "She squeals")) %>%
    distinct(scriptID, text) %>%
    select(text) %>%
    tail() %>%
    kable()
```

Depending on the visual design, we could have *one* example text for each highlighted word perhaps, visible when the user hovers, or several that randomly switch? We could also include film title information there.

## Joining the bigrams to other metadata about the scripts

```{r pronoun_imdb, dependson="pronoun_bigrams"}

mapping <- read_csv("full_mapping.csv") %>%
    rename(imdb = imdb_id)

gender <- read_csv("gender.csv") %>%
    filter(important == "1",
           gender != "NULL")

genre <- read_tsv("imdb-genre.tsv") %>%
    rename(imdb = imdb_id)

metadata <- read_tsv("imdb-meta-data-title.tsv") %>%
    select(imdb, year) ## could also get title here if we want later

pronoun_imdb <- pronoun_bigrams %>%
    left_join(mapping, by = c("scriptID" = "id")) 
```

## Differences by gender of writer

We can use the odds ratio to find words coming after "she" that are more likely to be used by female writers/directors/etc and those more likely to be used by male writers/directors/etc. Movies can have both, can have multiple writers and so forth, but the following analysis accounts for that.

```{r gender_ratio, dependson="pronoun_imdb", fig.width=7, fig.height=6}
gender_ratio <- pronoun_imdb %>%
    inner_join(gender, by = "imdb") %>% ## to get gender of writers, etc
    filter(word1 == "she",
           str_detect(word2, "[a-z]+"),
           nchar(word2) > 2) %>%
    group_by(word2) %>%
    mutate(total = n_distinct(scriptID)) %>%
    ungroup %>%
    count(gender, word2, total) %>%
    spread(gender, n, fill = 0) %>%
    mutate(female = (female + 1) / sum(female + 1),
           male = (male + 1) / sum(male + 1),
           ratio = female / male,
           logratio = log2(ratio)) %>%
    arrange(desc(logratio))

gender_ratio %>%
    filter(total > 50) %>%
    group_by(logratio < 0) %>%
    top_n(15, abs(logratio)) %>%
    ungroup() %>%
    mutate(word = reorder(word2, logratio)) %>%
    ggplot(aes(word, ratio - 1, color = logratio < 0)) +
    geom_segment(aes(x = word, xend = word,
                     y = 0, yend = ratio - 1), 
                 size = 1.1, alpha = 0.6) +
    geom_point(size = 3.5) +
    coord_flip() +
    labs(x = NULL, 
         y = "Relative use by women / men",
         title = "Words paired with 'she' from writers/directors/etc of different genders",
         subtitle = "From 4000+ films, verbs mentioned 50+ times\nWomen are more likely to portray women who write, fling, remember, wander\nMen are more likely to portray women who respond, emerge, believe, react") +
    scale_color_discrete(name = "", labels = c("More from women", "More from men")) +
    scale_y_continuous(breaks = c(-0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8),
                       labels = c("-60%", "-40%", "-20%", "Same", "+20%", "+40%", "+60%", "+80%"))
```

What are some words after "she" that have about the same likelihood to come from writers/directors/etc who are men and women?

```{r dependson="gender_ratio"}
gender_ratio %>%
    filter(total > 50) %>%
    arrange(abs(logratio))
```

Words like "still", "and", "drives", "tries", "brings", "listens"...

We could make another visualization in the same style as the first one, but I worry that users may be confused since they are communicating different information. Maybe they should visually look quite different, perhaps bars for this one if we do circles for the first one. We could choose to only show the extremes here.

## Change by year

We can use logistic regression modeling to find the words paired with "she" that have changed the fastest with time, either in a positive or negative direction.

```{r words_by_year, dependson="pronoun_imdb"}
words_by_year <- pronoun_imdb %>%
    inner_join(metadata, by = "imdb") %>% ## to get year
    filter(word1 == "she",
           str_detect(word2, "[a-z]+"),
           word2 != "look",
           nchar(word2) > 2,
           !is.na(year),
           year > 1980) %>%
    mutate(year = 5 * year %/% 5) %>%
    add_count(year) %>%
    rename(year_total = n) %>%
    add_count(word2) %>%
    rename(word_total = n) %>%
    filter(word_total > 50) %>%
    count(word2, year, year_total)

slopes <- words_by_year %>%
    nest(-word2) %>%
    mutate(models = map(data, ~ glm(cbind(n, year_total) ~ year, ., 
                                    family = "binomial"))) %>%
    unnest(map(models, tidy)) %>%
    filter(term == "year") %>%
    arrange(estimate)
```

```{r contractions, dependson="words_by_year", fig.width=7, fig.height=5}
words_by_year %>%
    inner_join(slopes %>%
                   top_n(10, estimate), 
               by = "word2") %>%
    ggplot(aes(year, n / year_total, color = word2)) +
    geom_line(alpha = 0.8, size = 1.3) +
    labs(x = NULL, y = "Word frequency",
         title = "Words paired with 'she' in script set directions",
         subtitle = "Many of the words increasing in frequency with time are negative contractions")
```

```{r increasing, dependson="words_by_year", fig.width=7, fig.height=5}
words_by_year %>%
    inner_join(slopes %>%
                   filter(str_detect(word2, "^[^’]*$")) %>%
                   top_n(10, estimate), 
               by = "word2") %>%
    ggplot(aes(year, n / year_total, color = word2)) +
    geom_line(alpha = 0.8, size = 1.3) +
    labs(x = NULL, y = "Word frequency",
         title = "Words paired with 'she' in script set directions",
         subtitle = "More recently, women are more likely to inspect, drift, steal, exhale,  breathe")
```

```{r decreasing, dependson="words_by_year", fig.width=7, fig.height=5}
words_by_year %>%
    inner_join(slopes %>%
                   top_n(-10, estimate), 
               by = "word2") %>%
    ggplot(aes(year, n / year_total, color = word2)) +
    geom_line(alpha = 0.8, size = 1.3) +
    labs(x = NULL, y = "Word frequency",
         title = "Words paired with 'she' in script set directions",
         subtitle = "In the past, women were more likely to jab, snuggle, fling, whirl, shriek")
```

I'm picturing an interactive version of this chart where the user can switch from increasing to decreasing, hover over the lines to see how they are changing. Maybe decide how many to see?

Let's check the words that change for "he", to compare. (This probably won't go into the final essay.)

```{r words_by_year_he, dependson="pronoun_imdb"}
words_by_year <- pronoun_imdb %>%
    inner_join(metadata, by = "imdb") %>% ## to get year
    filter(word1 == "he",
           str_detect(word2, "[a-z]+"),
           word2 != "look",
           nchar(word2) > 2,
           !is.na(year),
           year > 1980) %>%
    mutate(year = 5 * year %/% 5) %>%
    add_count(year) %>%
    rename(year_total = n) %>%
    add_count(word2) %>%
    rename(word_total = n) %>%
    filter(word_total > 100) %>%
    count(word2, year, year_total)

slopes <- words_by_year %>%
    nest(-word2) %>%
    mutate(models = map(data, ~ glm(cbind(n, year_total) ~ year, ., 
                                    family = "binomial"))) %>%
    unnest(map(models, tidy)) %>%
    filter(term == "year") %>%
    arrange(estimate)

# increasing for "he"
slopes %>% 
    top_n(10, estimate)

# decreasing for "he"
slopes %>% 
    top_n(-10, estimate)
```

So the change in contractions looks like it is more about how set directions are changing as a whole, not a difference for how women are portrayed. For the final modeling by year, I'll plan to filter contractions out.

Also, HA -- women are exhaling and men are inhaling.