### 2017-07-16 draft_post.Rmd
Working on just writers

---
title: "Women in Film"
author: "Julia Silge"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE, 
                      echo = FALSE, dpi = 360)
options(width=80, dplyr.width = 150)
library(ggplot2)
library(silgelib)
theme_set(theme_roboto())
```

How are women portrayed in film? It's complicated; the question might make you think of a male protagonist's romantic interest, someone in trouble who needs to be rescued, or maybe Wonder Woman.

One way to quantitatively measure how women are portrayed in film is to look at scene directions in film scripts. Scene directions are the part of a script that are not dialogue, describing what's happening on screen and telling actors how to move and how to deliver their lines. These scene directions communicate to the actors, producers, and director what story the script is working to tell.

```{r pronoun_bigrams, echo=FALSE}
library(tidyverse)
library(googlesheets)
library(tidytext)
library(stringr)

public_scripts <- gs_title("Public Script Sources")
script_sample <- public_scripts %>%
    gs_read("public_scripts", verbose = FALSE)

all_tsvs <- paste0("setDirections/", list.files(path = "setDirections/", pattern = ".tsv$"))
select_tsvs <- sort(paste0("setDirections/setDirections-", script_sample$script_id, ".tsv"))

sample_tsvs <- intersect(all_tsvs, select_tsvs)

all_lines <- sample_tsvs %>%
    map_df(~data_frame(lines = read_lines(.x)) %>%
               filter(str_detect(lines, "^[0-9]")) %>%
               separate(lines, c("line", "text"), sep = "\t") %>%
               mutate(scriptID = as.integer(str_extract(.x, "[\\d]+"))))

pronoun_bigrams <- all_lines %>%
    unnest_tokens(bigram, text, token = "ngrams", 
                  n = 2, collapse = FALSE) %>%
    separate(bigram, c("word1", "word2"), sep = " ") %>%
    filter(word1 %in% c("he", "she")) %>%
    select(scriptID, line, word1, word2) %>%
    filter(!str_detect(word2, "’|'"))
```

*Really rough mock-up of general idea for main viz:*

```{r pronoun_ratio, echo = FALSE, dependson="pronoun_bigrams", fig.width=5, fig.height=10}
pronoun_ratio <- pronoun_bigrams %>%
    count(word1, word2) %>%
    spread(word1, n, fill = 0) %>%
    mutate(total = he + she,
           he = (he + 1) / sum(he + 1),
           she = (she + 1) / sum(she + 1),
           logratio = log2(she / he)) %>%
    arrange(desc(logratio)) %>%
    filter(nchar(word2) > 1,
           !(word2 %in% c("door", "ad", "he", "1s", "lp",
                          "car", "himself", "pans", "dead")))

library(ggbeeswarm)
pronoun_ratio %>%
    top_n(800, total) %>%
    mutate(blank = "blank",
           word = reorder(word2, logratio)) %>%
    ggplot(aes(blank, logratio, label = word2)) +
    geom_beeswarm(color = "midnightblue", alpha = 0.4) +
    geom_text(family = "RobotoCondensed-Regular", size = 2.5, 
              vjust = 1.2, hjust = 1.2,
              check_overlap = TRUE) +
    theme_void()

pronoun_ratio %>%
    top_n(800, total) %>%
    mutate(freq = 2 ^ logratio) %>%
    write_csv("~/Desktop/pronoun_word_ratios.csv")
```

*Label all the outlier points, choose interesting other points to label. For all labeled points, have example text for interactive hover. For non-labeled points, show word on hover.*

Example text snippets for "she squeals":

```{r, echo = FALSE, dependson="pronoun_ratio"}
all_lines %>%
    filter(str_detect(text, "She squeals")) %>%
    distinct(scriptID, text) %>%
    select(text) %>%
    tail() %>%
    kable()
```

```{r pronoun_imdb, echo = FALSE, dependson="pronoun_bigrams"}

mapping <- read_csv("full_mapping.csv") %>%
    rename(imdb = imdb_id)

gender <- read_csv("gender.csv") %>%
    filter(important == "1",
           gender != "NULL",
           role == "writer")

genre <- read_tsv("imdb-genre.tsv") %>%
    rename(imdb = imdb_id)

metadata <- read_tsv("imdb-meta-data-title.tsv") %>%
    select(imdb, year) ## could also get title here if we want later

pronoun_imdb <- pronoun_bigrams %>%
    left_join(mapping, by = c("scriptID" = "id")) 
```

## Impact of writers

*Rough mockup for viz of writer viz*

```{r gender_ratio, echo = FALSE, dependson="pronoun_imdb", fig.width=8, fig.height=8}
library(viridis)

she_gender_ratio <- pronoun_imdb %>%
    inner_join(gender, by = "imdb") %>% ## to get gender of writers
    filter(word1 == "she",
           str_detect(word2, "[a-z]+"),
           nchar(word2) > 2) %>%
    group_by(word2) %>%
    mutate(total = n_distinct(scriptID)) %>%
    ungroup %>%
    count(gender, word2, total) %>%
    spread(gender, n, fill = 0) %>%
    mutate(female = (female + 1) / sum(female + 1),
           male = (male + 1) / sum(male + 1),
           she_ratio = female / male,
           she_logratio = log2(she_ratio)) %>%
    arrange(desc(she_logratio))

he_gender_ratio <- pronoun_imdb %>%
    inner_join(gender, by = "imdb") %>% ## to get gender of writers, etc
    filter(word1 == "he",
           str_detect(word2, "[a-z]+"),
           nchar(word2) > 2) %>%
    group_by(word2) %>%
    mutate(total = n_distinct(scriptID)) %>%
    ungroup %>%
    count(gender, word2, total) %>%
    spread(gender, n, fill = 0) %>%
    mutate(female = (female + 1) / sum(female + 1),
           male = (male + 1) / sum(male + 1),
           he_ratio = female / male,
           he_logratio = log2(he_ratio)) %>%
    arrange(desc(he_logratio))

she_gender_ratio %>%
    select(word2, she_ratio, she_logratio, she_total = total) %>%
    left_join(he_gender_ratio %>%
                  select(word2, he_ratio, he_logratio, he_total = total)) %>%
    top_n(300, she_total) %>%
    ggplot(aes(she_logratio, he_logratio, label = word2,
               size = (she_total + he_total),
               color = she_ratio / he_ratio)) +
    annotate("text", x = 0.8, y = 0.8,
             label = "Male and female writers\ndescribing women",
             family = "RobotoCondensed-Regular") +
    annotate("text", x = -0.8, y = -0.8,
             label = "Male and female writers\ndescribing men",
             family = "RobotoCondensed-Regular") +
     annotate("text", x = 0.8, y = -0.8,
             label = "Female writers describing women\nMale writers decribing men",
             family = "RobotoCondensed-Regular") +
     annotate("text", x = -0.8, y = 0.8,
             label = "Female writers describing men\nMale writers decribing women",
             family = "RobotoCondensed-Regular") +
   geom_vline(xintercept = 0, color = "gray50", lty = 2, size = 1.5) +
    geom_hline(yintercept = 0, color = "gray50", lty = 2, size = 1.5) +
    geom_text(family = "RobotoCondensed-Regular", size = 3, 
              vjust = 1.2, hjust = 1.2,
              check_overlap = TRUE) +
    geom_point(alpha = 0.5) +
    theme_void() +
    scale_y_continuous(limits = c(-1.1, 1.1)) +
    scale_x_continuous(limits = c(-1.1, 1.1)) +
    scale_color_viridis(end = 0.8) +
    theme(legend.position="none")

```

*Hover for word for all points*

Add bar plot for just how women are portrayed

## Word use changing with time

```{r changing_women, echo = FALSE, dependson="pronoun_imdb", fig.width=8, fig.height=6}
library(ggrepel)
words_by_year <- pronoun_imdb %>%
    inner_join(metadata, by = "imdb") %>% ## to get year
    filter(word1 == "she",
           str_detect(word2, "[a-z]+"),
           !(word2 %in% c("look", "doesn")),
           nchar(word2) > 2,
           !is.na(year),
           year > 1980) %>%
    mutate(year = 5 * year %/% 5) %>%
    add_count(year) %>%
    rename(year_total = n) %>%
    add_count(word2) %>%
    rename(word_total = n) %>%
    filter(word_total > 50) %>%
    count(word2, year, year_total)

slopes <- words_by_year %>%
    nest(-word2) %>%
    mutate(models = map(data, ~ glm(cbind(n, year_total) ~ year, ., 
                                    family = "binomial"))) %>%
    unnest(map(models, tidy)) %>%
    filter(term == "year") %>%
    arrange(estimate)

plot_words <- words_by_year %>%
    inner_join(slopes %>%
                   top_n(10, estimate), 
               by = "word2") %>%
    mutate(year = 10 * year %/% 10) %>%
    filter(year %in% c(1980, 2010)) %>%
    group_by(word2, year) %>%
    summarise(n = sum(n),
              total = sum(year_total)) %>%
    mutate(Change = "Increasing") %>%
    bind_rows(words_by_year %>%
                  inner_join(slopes %>%
                                 top_n(-10, estimate), 
                             by = "word2") %>%
                  mutate(year = 10 * year %/% 10) %>%
                  filter(year %in% c(1980, 2010)) %>%
                  group_by(word2, year) %>%
                  summarise(n = sum(n),
                            total = sum(year_total)) %>%
                  mutate(Change = "Decreasing")) %>% 
    ungroup

ggplot(plot_words, aes(year, n / total, color = word2)) +
    geom_line(size = 1.5, alpha = 0.5) +
    geom_point(size = 4) +
    geom_text_repel(data = subset(plot_words, year == 1980 & Change == "Decreasing"),
                  aes(x = year, y = n / total, label = word2),
                  color = "gray20", family = "RobotoCondensed-Regular",
                  point.padding = unit(0.4, "lines")) +
    geom_text_repel(data = subset(plot_words, year == 2010 & Change == "Increasing"),
                  aes(x = year, y = n / total, label = word2),
                  color = "gray20", family = "RobotoCondensed-Regular",
                  point.padding = unit(0.4, "lines")) +
    theme(legend.position="none") +
    facet_wrap(~Change, ncol = 2) +
    labs(x = NULL, y = "Word frequency",
         title = "Words paired with 'she' in script set directions")
```

```{r changing_men, echo = FALSE, dependson="pronoun_imdb", fig.width=8, fig.height=6}
library(ggrepel)
words_by_year <- pronoun_imdb %>%
    inner_join(metadata, by = "imdb") %>% ## to get year
    filter(word1 == "he",
           str_detect(word2, "[a-z]+"),
           !(word2 %in% c("look", "doesn", "were", 
                          "too", "see", "doing", "going")),
           nchar(word2) > 2,
           !is.na(year),
           year > 1980) %>%
    mutate(year = 5 * year %/% 5) %>%
    add_count(year) %>%
    rename(year_total = n) %>%
    add_count(word2) %>%
    rename(word_total = n) %>%
    filter(word_total > 50) %>%
    count(word2, year, year_total)

slopes <- words_by_year %>%
    nest(-word2) %>%
    mutate(models = map(data, ~ glm(cbind(n, year_total) ~ year, ., 
                                    family = "binomial"))) %>%
    unnest(map(models, tidy)) %>%
    filter(term == "year") %>%
    arrange(estimate)

plot_words <- words_by_year %>%
    inner_join(slopes %>%
                   top_n(10, estimate), 
               by = "word2") %>%
    mutate(year = 10 * year %/% 10) %>%
    filter(year %in% c(1980, 2010)) %>%
    group_by(word2, year) %>%
    summarise(n = sum(n),
              total = sum(year_total)) %>%
    mutate(Change = "Increasing") %>%
    bind_rows(words_by_year %>%
                  inner_join(slopes %>%
                                 top_n(-10, estimate), 
                             by = "word2") %>%
                  mutate(year = 10 * year %/% 10) %>%
                  filter(year %in% c(1980, 2010)) %>%
                  group_by(word2, year) %>%
                  summarise(n = sum(n),
                            total = sum(year_total)) %>%
                  mutate(Change = "Decreasing")) %>% 
    ungroup

ggplot(plot_words, aes(year, n / total, color = word2)) +
    geom_line(size = 1.5, alpha = 0.5) +
    geom_point(size = 4) +
    geom_text_repel(data = subset(plot_words, year == 1980 & Change == "Decreasing"),
                  aes(x = year, y = n / total, label = word2),
                  color = "gray20", family = "RobotoCondensed-Regular",
                  point.padding = unit(0.4, "lines")) +
    geom_text_repel(data = subset(plot_words, year == 2010 & Change == "Increasing"),
                  aes(x = year, y = n / total, label = word2),
                  color = "gray20", family = "RobotoCondensed-Regular",
                  point.padding = unit(0.4, "lines")) +
    theme(legend.position="none") +
    facet_wrap(~Change, ncol = 2) +
    labs(x = NULL, y = "Word frequency",
         title = "Words paired with 'he' in script set directions")
```

*interactively switch from men to women, increasing to decreasing*